Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-01-06 sources/bmc_template_submission/bmc_article.tex
intro

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{Dobzhansky maybe}.
A challenge for phylogenetic reconstruction
is assembling high-quality homology hypotheses.
While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{BLA}, creating challenges for
reconstructing the shared evolutionary history of taxa separated phylogenetically
at deep scales \cite{BLA}.

While phylogenomics tackles this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. It does not contribute to
democratization of research yet.

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that can be used for phylogenetic
reconstruction at many scales.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
How much of this is available on OpenTree??

@@

Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
integration of data into phylogenies, which can be addressed with a unified
taxonomy for name standardization. The Open Tree of Life project (OpenTree)
constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic
data from various databases \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) molecular database GenBank
\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
for connecting data from any biological database that has been integrated to it.

@@

>>>>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{dobzhansky1973nothing}.
A challenge for phylogenetic reconstruction
is assembling high-quality homology hypotheses.
While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
creating challenges for phylogenetic reconstruction.

While phylogenomics tackles this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide}, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. It does not contribute to
democratization of research yet \cite{li2017enabling}.

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that can be used for phylogenetic
reconstruction at many scales.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
The Open Tree of Life project (OpenTree)
constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. Meaning that there is a lot of information

@@

Make emphasis on these really great studies (and pipelines) that do not rely on
genomic data, and that there is a lot of single locus data that can be analysed

@@

Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
integration of data into phylogenies, which can be addressed with a unified
taxonomy for name standardization.

To achieve this, OpenTree unifies taxonomic
data from various databases \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) molecular database GenBank
\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
for connecting data from any biological database that has been integrated to it.
